Type theories as quotient inductive-recursive types

### 2025-02-21 tex/main.tex
Polish proofs until Prop 3.16 in Appendix

\propCoherence*
\begin{proof}
  For term substitution, the claim follows by \cref{prop:correctness-1}: if
  $\cc : \sigma = \tau$ then
  \begin{align*}
    \subTm{t}{\sigma} & = \sub{t}{\sigma} && \text{by \cref{prop:correctness-1}} \\
                      & = \sub{t}{\tau} && \text{by congruence and $\cc$} \\
                      & = \subTm{t}{\tau} && \text{by \cref{prop:correctness-1}}
  \end{align*}
  For type substitution, by \cref{re:coherence-proof}, we only have to consider the following cases
  \begin{align*}
    \subTy{A}{\sigma; (\tau, t)}    & = \subTy{A}{(\sigma;\tau), \sub{t}{\sigma}},
                                     & \subTy{A}{\sigma}               & = \subTy{A}{\emptyctx},
                                     & \subTy{A}{\sigma}               & = \subTy{A}{\pi_1\sigma, \pi_2\sigma}
  \end{align*}
  while other cases hold strictly.
  Again, for each case, we prove them by induction on $A$ and consider the case $A \defeq \El\,u$ only which is added to the definition of $[\blank]_{\Ty}\blank$ in this section. 
  However, all cases follow easily from \cref{prop:correctness-1}.
  For example, consider the first case:
  \begin{align*}
    \subTy{(\El\,u)}{\sigma; (\tau, t)} & \judgeq \El(\subTm{u}{\sigma; (\tau, t)}) && \text{by \eqref{eq:def-type-subst-7}} \\
%                                      & = \El( \sub{u}{\sigma; (\tau, t)})            && \text{by \cref{prop:correctness-1}} \\
%                                      & = \El( \sub{u}{\sigma; \tau, \sub{t}{\sigma}})     && \text{by $\mathsf{concat}$} \\
%                                      & = \El( \subTm{u}{\sigma; \tau, \sub{t}{\sigma}}) && \text{by \cref{prop:correctness-1}} \\
                                      & = \El( \subTm{u}{\sigma; \tau, \sub{t}{\sigma}}) && \text{by coherence of $\subTm{\blank}{\blank}$} \\
                                      & \judgeq \subTy{\left(\El\,u\right)}{\sigma; \tau, \sub{t}{\sigma}} && \text{by \eqref{eq:def-type-subst-7} }
  \end{align*}
  The other two cases are dealt with in exactly the same way.
\end{proof}

@@

\lemIdentities*
\begin{proof}
  We have
  \[
    \idS^+ \judgeq ((\pi_1\idS; \idS), \pi_2 \idS) = (\pi_1\idS, \pi_2 \idS) = \idS
  \]
  For commuting with composition, first note that $\pi_1(\sigma; \tau) = \sigma; \pi_1 \tau$ by $\mathsf{concat}$. In particular, we have $\sigma; \pi_1 \idS = \sigma^+; \pi_1\idS$. Now
  \begin{align*}
    (\sigma; \tau)^+ &= (\pi_1\idS; (\sigma; \tau), \pi_2 \idS) \\
                     &= ((\pi_1\idS; \sigma); \tau, \pi_2 \idS) \\
                     &= ((\sigma^+; \pi_1\idS); \tau, \sub{\pi_2 \idS}{\sigma^+}) \\
                     &= (\sigma^+; (\pi_1\idS; \tau), \pi_2 \idS) \\
                     &= \sigma^+; \tau^+
  \end{align*}
  as required.
\end{proof}

>>>>

\propCoherence*
\begin{proof}
  For term substitution, the claim follows by \cref{prop:correctness-1}: if
  $\cc : \sigma = \tau$ then
  \begin{align*}
    \subTm{t}{\sigma} & = \sub{t}{\sigma} && \text{by \cref{prop:correctness-1}} \\
                      & = \sub{t}{\tau} && \text{by congruence and $\cc$} \\
                      & = \subTm{t}{\tau} && \text{by \cref{prop:correctness-1}}
  \end{align*}
  For type substitution, by \cref{re:coherence-proof}, we only have to consider the following cases
  \begin{align*}
    \subTy{A}{\sigma; (\tau, t)}    & = \subTy{A}{(\sigma;\tau), \sub{t}{\sigma}},
                                     & \subTy{A}{\sigma}               & = \subTy{A}{\emptyctx},
                                     & \subTy{A}{\sigma}               & = \subTy{A}{\pi_1\sigma, \pi_2\sigma}
  \end{align*}
  while other cases hold strictly.
  Again, for each case, we prove them by induction on $A$ and consider the case $A \defeq \El\,u$ only which is added to the definition of $[\blank]_{\Ty}\blank$ in this section. 
  However, all cases follow easily from \cref{prop:correctness-1}.
  For example, consider the first case:
  \begin{align*}
    \subTy{(\El\,u)}{\sigma; (\tau, t)} & \judgeq \El\,(\subTm{u}{\sigma; (\tau, t)}) && \text{by \eqref{eq:def-type-subst-7}} \\
%                                      & = \El( \sub{u}{\sigma; (\tau, t)})            && \text{by \cref{prop:correctness-1}} \\
%                                      & = \El( \sub{u}{\sigma; \tau, \sub{t}{\sigma}})     && \text{by $\mathsf{concat}$} \\
%                                      & = \El( \subTm{u}{\sigma; \tau, \sub{t}{\sigma}}) && \text{by \cref{prop:correctness-1}} \\
                                      & = \El\,( \subTm{u}{\sigma; \tau, \sub{t}{\sigma}}) && \text{by coherence of $\subTm{\blank}{\blank}$} \\
                                      & \judgeq \subTy{\left(\El\,u\right)}{\sigma; \tau, \sub{t}{\sigma}} && \text{by \eqref{eq:def-type-subst-7} }
  \end{align*}
  The other two cases are dealt with in exactly the same way.
\end{proof}

@@

\lemIdentities*
\begin{proof}
  By the identity law $\mathsf{idr}$ and the $\eta$-law $\pi\eta$, we have
  \[
    \idS^+ \judgeq ((\pi_1\idS; \idS), \pi_2 \idS) = (\pi_1\idS, \pi_2 \idS) = \idS.
  \]
  For commuting with composition, first note that $\pi_1(\sigma; \tau) = \sigma; \pi_1 \tau$ holds similarly to \eqref{eq:pi2-comp-proof} and, in particular, we have $\pi_1 \idS ; \sigma = \sigma^+; \pi_1\idS$.
  Also note that $\pi_2\idS = \sub{\pi_2\idS}{\sigma^+}$ by \eqref{eq:pi2-comp-proof}.
  Now
  \begin{align*}
    (\sigma; \tau)^+ & \judgeq (\pi_1\idS; (\sigma; \tau)), \pi_2 \idS && \text{by definition} \\
                     & = ((\pi_1\idS; \sigma); \tau), \pi_2 \idS       && \text{by ;-$\mathsf{assoc}$}\\
                     & = ((\sigma^+; \pi_1\idS); \tau), \sub{\pi_2 \idS}{\sigma^+} && \text{by $\pi_1\idS; \sigma = \sigma^+ ; \pi_1\idS$ and $\pi_2\idS = \sub{\pi_2\idS}{\sigma^+}$}\\
                     & = (\sigma^+; (\pi_1\idS; \tau)) , \sub{\pi_2\idS}{\sigma^+}  && \text{by ;-$\mathsf{assoc}$} \\
                     & = \sigma^+; ((\pi_1\idS ; \tau) , \pi_2\idS) && \text{by $\mathsf{concat}^{-1}$} \\
                     & \judgeq \sigma^+; \tau^+ && \text{by definition}
  \end{align*}
  as required.
\end{proof}
